Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-12-14 demystify.tex
updates

% Below are some tips to approach your LoR writers:
% \begin{itemize}
% \item \textbf{Ask in advance}: You should ask for LoRs \emph{at least a month} before the deadline. People are busy with their own work and likely have many other students asking for LoRs. Also, a strong LoR takes time to write, so you should give them enough time to write a strong letter for you.

% \item \textbf{Help your writers}: You should provide your writers with details about the programs you are applying to, the deadlines, and any other information they need to know.  You should also provide them with details about your research experience and potential so that they can write a strong letter for you.
% \item \textbf{Follow up}: After you ask for LoRs, you should follow up with your writers to make sure they have submitted the letters on time.  You should also thank them for writing the letters for you.
% \item \textbf{Waive your right}: When you ask someone to write a letter for you, you will be asked if you want to waive your right to see the letter.  You should always waive your right.  This shows that you trust your writers and that you are not trying to twist their words.
% \item \textbf{Remind your writers}: After you enter your writers' information in the application system, you should tell your writers about that and let them know they will soon receive an email from the university to submit their letters.  You should also remind them when you submit your application and ask them to submit their letters on time if they haven't done so.
% \item \textbf{Help your writers}: To help your writers write a strong letter for you, you should provide them with details about your research experience and potential.  You should also provide them with a draft of your SOP so that they can see what you are saying about yourself and complement that with their own perspective.
% \item \textbf{Thank your writers}: After your writers have submitted their letters, you should thank them for writing the letters for you.  You should also let them know the outcome of your applications.
% \item \textbf{Ask for feedback}: After you have been admitted or rejected, you should ask your writers for feedback on your application.  This will help you improve your application for future applications.
% \item \textbf{Stay in touch}: After you have been admitted or rejected, you should stay in touch with your writers.  This will help you build a relationship with them and will help you in the future.
% \end{itemize}

@@

I will also let the student know if I cannot write a strong letter for them (e.g., I don't know them that well), and suggest they find someone else.  If they insist, then I will write for them. While I try to say something positive, e.g., the student is hardworking, or that I heard they are working on a research project, etc, the letter will still be short and weak (\autoref{sec:generic-letters}). Usually it takes me about 5--10 minutes to write such a letter.

For a student that I know well, I will write them a strong letter, which takes a lot longer as it will be personalized.  While I have a general template for LoR, it still can take me an hour or more to write it.  I often ask the students to provide information (\autoref{sec:help-your-LOR-writers}) and what they think I should highlight in the letter. They can also provide me their SOP (if they already have written one) so I can complement what they say with my own perspective.  I often do not share what I write with the students, just to keep it more genuine and honest.

>>>>

Below are some tips to approach your LoR writers:
\begin{itemize}
\item \textbf{Ask in advance} You should ask for LoRs \emph{at least a month} before the deadline. People are busy and writing a strong LoR takes time (\autoref{sec:lor-writing}), so you should give them enough time to write a strong letter for you.

\item \textbf{Waive your right (\autoref{sec:waive-right})} You should always waive your right to see the letter. This shows that you trust your writers and that you are not trying to twist their words.

\item \textbf{Help your writers (\autoref{sec:help-your-LOR-writers})}: You should tell your writers the programs you are applying to, their deadlines, etc.  You can also share your SOP with them and other details about your research experience and potential.

\item \textbf{Ask for feedback} If the writer is very close to you, you can ask them for feedback on your application, e.g., your SOP.

\item \textbf{Follow up and Stay in touch} Follow up with your writers to make sure they have submitted the letters on time.  After your writers have submitted all of their letters, \emph{thank them}!  Finally, let them know the outcome of your applications and stay in touch with them.  This will help you build a relationship with them and you might need them to write for you again in the future.
\end{itemize}

@@

I will also let the student know if I cannot write a strong letter for them (e.g., I don't know them that well), and suggest they find someone else.  If they insist, then I will write for them. While I try to say something positive, e.g., the student is hardworking, or that I heard they are working on a research project, etc, the letter will still be short and weak (\autoref{sec:generic-letters}). Usually it takes me about \emph{5--10} minutes to write such a letter.

Strong letters take a lot longer as it will be personalized.  While I have a general template for LoR, it still can take me \emph{an hour or more} to write it.  I often ask the students to provide information (\autoref{sec:help-your-LOR-writers}) and what they think I should highlight in the letter. They can also provide me their SOP (if they already have written one) so I can complement what they say with my own perspective.  I often do not share what I write with the students, just to keep it more genuine and honest.

My letter always has the university letterhead, my signature, and is signed and dated.  While using a letterhead or signing the letter is not required (I never paid much attention to these when reading LoRs), it makes the letter more official and professional.  
This website has the template of the letter with letterhead and signature: \url{https://www.overleaf.com/read/xzyrxkdjfxsp#2a1f9e}.
